% ============================================
% ABSTRACT & INDEX TERMS
% ============================================
% 👤 PERSON 1: Lead / Integrator
% 
% Instructions:
% - Write a concise abstract (150-200 words)
% - Focus on: problem, solution (ADRDS), key innovation, demo value
% - Mention the interactive nature of the demo
% - NO symbols, math, or footnotes in abstract
% - Keep Index Terms relevant to ransomware, ML, behavioral analysis
% ============================================

\begin{abstract}

Over the past few years, ransomware has grown into one of the most damaging threats facing organizations and individuals alike. When attackers encrypt critical files and demand payment, the consequences often include massive financial losses, disrupted operations, and eroded trust. Unfortunately, conventional antivirus tools that rely on known signatures frequently miss newer ransomware strains, especially those using obfuscation or appearing for the first time as zero-day threats.

In this paper, we introduce ADRDS, an AI-Driven Ransomware Detection and Response Simulator built around behavioral analysis rather than static pattern matching. Our system recreates ransomware-like behaviors in a safe, controlled setting, capturing telltale signs such as sudden spikes in file entropy, unusual sequences of API calls, and bursts of file read-write activity. We then feed these behavioral fingerprints into a hybrid machine learning pipeline. On one side, supervised classifiers trained on established datasets like BigMalware and EMBER recognize familiar attack patterns. On the other, an unsupervised LSTM Autoencoder flags anomalies that could signal brand-new variants the supervised models have never encountered.

By combining scores from both approaches into a single risk metric, ADRDS catches threats early and with greater confidence. The platform also features automated response simulations and a real-time dashboard so users can watch detection unfold step by step. Our experiments show that this hybrid strategy delivers strong, reliable detection without sacrificing safety or scalability, making ADRDS a valuable tool for security research, hands-on training, and building smarter cyber defenses.

\end{abstract}

\begin{IEEEkeywords}
ransomware detection, behavioral analysis, machine learning, security demonstration, 
threat modeling, intrusion detection
\end{IEEEkeywords}
