% ============================================
% EXPERIMENTAL SETUP
% ============================================
% 👤 PERSON 5: ML & Evaluation Lead
% ============================================

\section{Experimental Setup}
\label{sec:experiments}

To ensure the scientific validity of our findings, we established a rigorous experimental framework. This section details the dataset composition, feature engineering process, and the specific hyperparameters used for model training.

\subsection{Datasets}

We evaluate ADRDS using a hybrid dataset constructed from both simulated behaviors and real-world malware traces, consisting of:
\begin{itemize}
    \item \textbf{Ransomware samples:} We utilized 500 verified ransomware executables sourced from the VirusShare repository, covering five distinct families including WannaCry, LockBit, and Ryuk to ensure the model learns diverse encryption behaviors across different strains.
    
    \item \textbf{Benign samples:} To represent normal activity, we collected 500 benign samples comprising standard Windows system files, office productivity software, and web browsers, ensuring the model learns to distinguish malicious encryption from legitimate high-load system usage.
    
    \item \textbf{Total behavioral sequences:} The final dataset comprises 1,000 fully labeled behavioral sequences, balanced evenly between malicious and benign classes to prevent class imbalance bias during the supervised learning phase of our model training.
\end{itemize}

Table~\ref{tab:dataset} summarizes the statistical characteristics of the dataset used for validation.

% TABLE 1: DATASET CHARACTERISTICS
\begin{table}[htbp]
\caption{Dataset Characteristics}
\label{tab:dataset}
\centering
\begin{tabular}{|l|r|}
\hline
\textbf{Characteristic} & \textbf{Value} \\
\hline
Ransomware samples & 500 \\
Benign samples & 500 \\
Distinct Families & 5 (WannaCry, LockBit, Ryuk, etc.) \\
Total Features Extracted & 42 \\
Avg. Sequence Length & 1,250 events \\
\hline
\end{tabular}
\end{table}

\subsection{Feature Engineering}

We extract 42 distinct behavioral features from the raw API logs and system event sequences, focusing on:
\begin{itemize}
    \item \textbf{File operation statistics:} We calculate rates of read, write, and delete operations within sliding windows, specifically monitoring for Shannon entropy spikes which indicate the rapid transformation of structured data into encrypted pseudo-random ciphertext.
    
    \item \textbf{Process behavior patterns:} We track the frequency of child process spawning and attempts to inject code into remote threads, as these are strong indicators of privilege escalation and persistence mechanisms used by modern malware.
    
    \item \textbf{Registry modification counts:} We quantify changes to sensitive registry hives, particularly the "Run" keys and service configurations, which ransomware commonly modifies to ensure it survives system reboots and maintains control over the victim machine.
    
    \item \textbf{Network communication features:} We analyze the frequency of outbound connection attempts and total data volume transferred, identifying potential Command-and-Control (C2) beacons used for exchanging encryption keys or receiving instructions from attacker servers.
    
    \item \textbf{Cryptographic API usage:} We monitor the invocation rate of specific Windows Cryptographic APIs, such as \texttt{CryptEncrypt} and \texttt{CryptGenKey}, serving as a high-fidelity signal for the start of the encryption phase.
\end{itemize}

Feature normalization applies z-score standardization to scale all inputs to a mean of zero. Feature selection uses mutual information scoring to retain the top 20 most discriminative features for efficiency.

\subsection{Machine Learning Models}

We evaluate three distinct classifier types to determine the optimal balance between accuracy and speed:
\begin{itemize}
    \item \textbf{Random Forest:} We configured an ensemble of 100 decision trees with a maximum depth of 20, optimized to capture non-linear feature interactions while preventing the model from overfitting to specific training noise.
    
    \item \textbf{Support Vector Machine (SVM):} We utilized a Radial Basis Function (RBF) kernel with a regularization parameter $C=1.0$ and gamma set to 'scale', allowing the model to find complex decision boundaries in high-dimensional feature space.
    
    \item \textbf{Neural Network:} We implemented a Multi-Layer Perceptron (MLP) with two hidden layers of 64 and 32 neurons respectively, using ReLU activation functions to capture deep latent patterns in the sequential behavioral data.
\end{itemize}

\subsection{Evaluation Methodology}

\textbf{Training/Testing:} We employed an 80/20 stratified split, reserving 20\% of data for final testing. During training, we used 5-fold cross-validation to ensure our accuracy metrics are robust and not artifacts of data selection.

\textbf{Metrics:} We assess performance using Accuracy, Precision, Recall, F1-score, and ROC-AUC. We also measure detection latency in milliseconds to ensure the system is fast enough for the real-time interactive demo requirement.

\textbf{Baseline Comparison:} We compare our behavioral approach against a standard signature-based detection baseline (ClamAV), demonstrating the superior ability of ADRDS to detect obfuscated ransomware variants that easily bypass static file scanning methods.
