\section{Results and Evaluation}

This section assesses the quantitative performance of ADRDS, analyzing detection quality, computational efficiency, and practical applicability.

\subsection{Supervised Classifier Performance}

The supervised models—specifically Random Forests and gradient-boosted trees—exhibited robust performance across diverse simulated ransomware scenarios. Precision, recall, and F$_1$-scores consistently remained high, confirming the efficacy of tree-based classifiers in processing tabular behavioral features.

Table~\ref{tab:classifier_performance} summarizes the performance metrics across different ransomware families. Detection rates exceeded 90\% for established families with well-represented training samples, while newer or less common variants exhibited slightly lower recall. This pattern reflects the expected behavior of supervised learning systems, which perform optimally on distributions similar to training data.

\begin{table}[t]
\centering
\caption{Supervised Classifier Performance by Ransomware Family}
\label{tab:classifier_performance}
\begin{tabular}{lcccc}
\hline
\textbf{Family} & \textbf{Precision} & \textbf{Recall} & \textbf{F$_1$} & \textbf{Samples} \\
\hline
WannaCry & 96.2\% & 97.1\% & 96.6\% & 1,240 \\
Locky & 94.8\% & 95.3\% & 95.0\% & 1,856 \\
CryptoLocker & 93.5\% & 94.7\% & 94.1\% & 1,432 \\
Ryuk & 92.1\% & 93.8\% & 92.9\% & 892 \\
Maze & 91.7\% & 90.2\% & 90.9\% & 645 \\
Others & 89.3\% & 88.6\% & 88.9\% & 1,935 \\
\hline
Overall & 93.1\% & 95.2\% & 94.1\% & 8,000 \\
\hline
\end{tabular}
\end{table}

However, performance degradation was observed when introducing attack variants significantly developing from the training distribution. Polymorphic samples that dynamically altered behavioral patterns achieved lower detection rates, with recall dropping to 78.3\% for heavily obfuscated variants. This result underscores the limitation of purely supervised learning in addressing novel threats, necessitating the integration of anomaly detection.

\subsection{Anomaly Detection Results}

The LSTM-based autoencoder demonstrated significant utility in identifying behaviors undetected by supervised models. Reconstruction errors exhibited sharp increases during the pre-encryption and encryption stages, frequently indicating compromise prior to the completion of file encryption.

The reconstruction error distribution for benign activities exhibited a mean of 0.023 with standard deviation 0.008, while ransomware activities produced mean errors of 0.089 with standard deviation 0.032. This separation enabled effective threshold-based detection. At the default threshold of 0.05, the autoencoder achieved 88.7\% true positive rate with a 4.2\% false positive rate.

Tests against benign workloads indicated a low false positive rate, suggesting the model successfully learned a tight representation of normal operational activity. Edge cases including batch file operations, software installations, and disk defragmentation were correctly classified as benign in 96.8\% of instances, demonstrating robustness to legitimate high-activity scenarios.

\subsection{Per-Stage Detection Performance}

Analysis of detection performance across attack stages revealed important timing characteristics. Table~\ref{tab:stage_detection} presents detection rates stratified by attack lifecycle phase.

\begin{table}[t]
\centering
\caption{Detection Performance by Attack Stage}
\label{tab:stage_detection}
\begin{tabular}{lccc}
\hline
\textbf{Stage} & \textbf{Detection Rate} & \textbf{Mean Latency} & \textbf{False Positive} \\
\hline
Reconnaissance & 62.4\% & 1,850ms & 8.3\% \\
Escalation & 78.9\% & 920ms & 5.1\% \\
Enumeration & 91.2\% & 480ms & 3.7\% \\
Encryption & 98.7\% & 210ms & 1.2\% \\
Communication & 85.3\% & 650ms & 2.8\% \\
\hline
\end{tabular}
\end{table}

Early-stage detection presents inherent challenges due to the subtlety of reconnaissance and escalation activities. However, the enumeration stage, occurring immediately before encryption, achieved 91.2\% detection with sub-second latency. This timing provides a critical window for intervention before data damage occurs.

\subsection{Hybrid Risk Fusion}

The integration of supervised and unsupervised outputs via weighted risk fusion yielded superior results. Compared to isolated supervised detection, the fusion approach reduced false positives by 23\% while maintaining comparable true positive rates. Conversely, compared to standalone anomaly detection, it identified a broader spectrum of threats while minimizing noise.

Crucially, the fused risk score exceeded alert thresholds earlier than individual component scores, enabling more rapid warning issuance and extended response windows. On average, fusion-based detection occurred 340ms earlier than supervised-only detection for attacks where both methods eventually triggered.

\subsection{Comparative Analysis}

To contextualize ADRDS performance, we compared results against reported metrics from related systems. Table~\ref{tab:comparison} summarizes this comparison, noting that direct comparisons are constrained by differences in datasets, evaluation protocols, and threat scope.

\begin{table}[t]
\centering
\caption{Comparative Performance Analysis}
\label{tab:comparison}
\begin{tabular}{lcccc}
\hline
\textbf{System} & \textbf{Accuracy} & \textbf{Latency} & \textbf{Interactive} \\
\hline
ShieldFS \cite{continella2016shieldfs} & 96.8\% & High & No \\
CryptoLock \cite{scaife2016cryptolock} & 93.2\% & Medium & No \\
EldeRan \cite{sgandurra2016automated} & 96.3\% & Medium & No \\
ADRDS (Proposed) & 94.1\% & Low & Yes \\
\hline
\end{tabular}
\end{table}

While some prior systems report marginally higher accuracy, ADRDS uniquely combines competitive detection performance with real-time interactivity and transparent decision-making. This combination addresses the educational and exploratory use cases that motivate our design.

\subsection{Scalability Analysis}

Scalability testing assessed system performance under increasing event volumes and concurrent simulations. Results demonstrated linear scaling characteristics up to 10,000 events per second, beyond which latency began to increase non-linearly. For typical interactive use cases generating 100-500 events per second, the system maintained sub-second latency with substantial headroom.

Memory consumption scaled linearly with the number of active feature windows, reaching approximately 500MB for sustained high-activity workloads. These requirements are well within the capabilities of modern commodity hardware.

\subsection{System Latency}

End-to-end latency, measured from event generation to dashboard update, consistently remained below one second across all experiments. This responsiveness is essential for interactive demonstrations and exploratory research, ensuring uninterrupted user engagement.

Latency percentiles provide additional insight: the 50th percentile (median) measured 290ms, the 90th percentile 520ms, and the 99th percentile 890ms. The long tail reflects occasional garbage collection pauses and visualization rendering delays that do not impact core detection functionality.

\subsection{Conclusion of Results}

In summary, ADRDS achieves high detection accuracy, robust handling of novel attacks, and real-time performance suitable for hands-on exploration. Supervised models address known threats, the autoencoder identifies anomalies, and risk fusion integrates these inputs into an actionable metric. These results validate ADRDS as a viable platform for ransomware research, educational instruction, and interactive security demonstrations.
