\section{System Overview}

ADRDS is designed as a modular, interactive system bridging abstract detection research with empirical experimentation. As illustrated in Fig.~\ref{fig:architecture}, the architecture comprises three integrated components: a behavioral simulator, a machine learning detection engine, and an interactive visualization layer. These components facilitate the simulation of ransomware activity, real-time detection monitoring, and the exploration of parameter sensitivity.

\subsection{System Architecture}

\subsubsection{Behavioral Simulator}
The behavioral simulator functions as the core of ADRDS, generating realistic ransomware behavior sequences without the execution of live malware. It implements a configurable state machine modeling the attack lifecycle—including reconnaissance, privilege escalation, persistence, encryption, and optional command-and-control communication. Each event corresponds to observable host-level activity, such as file operations, process creation, and resource utilization spikes. Parameters including encryption rate, file targeting strategy, and execution intensity are adjustable, allowing for the emulation of diverse ransomware families and tactics.

\subsubsection{Machine Learning Detector}
The detection engine processes structured features extracted from simulator output, applying trained machine learning models to classify activity as benign or malicious. The system supports multiple classifiers, including Random Forests \cite{breiman2001random}, Support Vector Machines \cite{takeuchi2018detecting}, and neural networks, enabling performance comparison under varying conditions. A critical feature is the adjustable confidence threshold, which allows users to modulate detection sensitivity and analyze the trade-off between false positive rates and detection capabilities.

\subsubsection{Interactive Visualization Layer}
The visualization layer transforms raw telemetry into interpretable insights. It renders attack timelines, feature histograms, and risk indicators in real time. The dashboard interface permits the modification of simulation settings and detection thresholds, reflecting the immediate impact of these adjustments without requiring experiment restarts.

\subsection{Data Flow}

ADRDS operates via a closed-loop pipeline ensuring synchronization between simulation, detection, and visualization. The process initiates with the behavioral simulator generating timestamped event streams corresponding to each attack stage. Subsequently, a feature extraction module converts these raw events into structured feature vectors, aggregating stage context with behavioral metrics. The machine learning detector then evaluates these vectors, producing classification outputs and confidence scores. Finally, the visualization layer renders the results, attack progress, and performance metrics in real time. User interactions—such as parameter adjustments and threshold modifications—are fed back into the simulator, closing the control loop.

This architecture establishes ADRDS as a dual-purpose platform, suitable for both demonstration and research. The modular design ensures that individual components can be replaced or extended, such as incorporating new models or feature extractors, without disrupting the overall pipeline.

% Figure placement
\begin{figure}[t]
\centering
\scalebox{0.75}{
\begin{tikzpicture}[
    node distance=0.6cm and 1cm,
    box/.style={rectangle, draw, fill=blue!10, text width=2.5cm, align=center, minimum height=0.7cm, font=\footnotesize},
    arrow/.style={-{Stealth[length=1.5mm]}, thick},
    feedback/.style={-{Stealth[length=1.5mm]}, thick, dashed, red}
]

% Nodes
\node[box] (simulator) {Behavioral Simulator};
\node[box, below=of simulator] (events) {Event Generation};
\node[box, below=of events] (features) {Feature Extraction};
\node[box, below=of features] (supervised) {Supervised Classifier};
\node[box, right=of supervised] (unsupervised) {Unsupervised Autoencoder};
\node[box, below=of supervised, xshift=1.25cm] (fusion) {Risk Fusion};
\node[box, below=of fusion] (viz) {Visualization \& Alerts};
\node[box, left=of viz, xshift=-0.4cm] (user) {User Interaction};

% Arrows
\draw[arrow] (simulator) -- (events);
\draw[arrow] (events) -- (features);
\draw[arrow] (features) -- (supervised);
\draw[arrow] (features) -| (unsupervised);
\draw[arrow] (supervised) -- (fusion);
\draw[arrow] (unsupervised) -- (fusion);
\draw[arrow] (fusion) -- (viz);
\draw[arrow] (viz) -- (user);
\draw[feedback] (user) |- ([yshift=0.25cm]simulator.north) -- (simulator);

\end{tikzpicture}
}
\caption{ADRDS system architecture illustrating data flow between the behavioral simulator, machine learning-based detector, and interactive visualization components. The dashed line represents the feedback loop.}
\label{fig:architecture}
\end{figure}
